### 2014-11-06 writedown/period.noneq.msm/coverletter/letter-jctc.tex
add keywords, some revises from CS

\documentclass{letter}

\usepackage{graphicx}

\renewcommand{\baselinestretch}{1.0}
% \setlength{\footskip}{0cm}
%\addtolength{\textheight}{10cm}
% \sloppy

\address{Han Wang\\
  Zuse Institute Berlin \\
  Takustrasse 7\\
  14195 Berlin, Germany 
}
\signature{Han Wang}

% \def\urltilda{\kern -.15em\lower .7ex\hbox{\~{}}\kern .04em}
% \def\urlunder{\lower .1ex\hbox{\_{}}\kern .04em}
% \def\urldot{\kern -.10em.\kern -.10em}
% \def\urlhttp{http\kern -.10em\lower -.1ex\hbox{:}\kern -.12em\lower 0ex\hbox{/}\kern -.18em\lower 0ex\hbox{/}}
% \parindent0ex

\begin{document}

\begin{letter}{
    The Editors of\\
    Journal of Chemical Theory and Computation\\
    \vskip 1cm
  }
% \begin{picture}(0,0)
%   \put(-30,22){
%     \parbox{0.25\textwidth}{
% %	\includegraphics[scale=0.4]{logo-icp-notext.pdf}
%     }
%   }
%   \put(264,10){
%     \parbox{0.25\textwidth}{
% 	\includegraphics[scale=0.28]{Fub-logo.eps}
%     }
%   }
% \end{picture}

\opening{Dear Editors,\vskip .5cm}
Hereby we would like to submit our article: 
\textit{``Building Markov State Models for Periodically Driven Non-Equilibrium Systems''
} by H. Wang and C. Sch\"utte to the
\textit{Journal of Chemical Theory and Computation}, as a regular article.

In this paper, we demonstrate how to build the Markov state model
(MSM) for the non-equilibrium systems under a periodic external
driving.  The MSM reproduces the essential long timescale
conformational dynamics of the original system, and at the same time
it does not require the ergodicity of each single molecular
trajectory.  Therefore, it does not only describes, but also provides
more insightful understanding of the non-equilibrium system than the
standard analyses of the molecular trajectories.  Our methods are
tested and validated by an example non-equilibrium system: The alanine
dipeptide under periodic electric field.

To the knowledge of the authors, this work is the first attempt of
building MSM for non-equilibrium systems under periodic external
driving. The significance of our method is that it does not assume
reversibility of the dynamics, as it is mostly done in building MSM
for equilibrium systems.

As experts who can judge the validity of the work we suggest the
following names:
\begin{itemize}\itemsep -1pt
\item Vijay Pande, Stanford University, pande@stanford.edu
\item John Chodera, Memorial Sloan Kettering Cancer Center, choderaj@mskcc.org
\item Giovanni Ciccotti,  University of Rome La Sapienza, Giovanni.Ciccotti@roma1.infn.it
\end{itemize}

\vskip 1cm

Best regards,\\
Han Wang

\end{letter} \end{document}